### 2022-08-02 main/chapter_2ljets.tex
wrap up signal theory variations

\paragraph{Cross-section variations}
\label{sec:2ljets_xsec_variations}
We scale all signal models to cross-sections calculated with the
\textsc{Resummino} package at NLO accuracy in QCD with NLL
(next-to-leading logarithmic) accuracy in the gaugino pair production of our
signal initial states~\cite{Fuks:2012qx, Fuks:2013vua}.
These cross-sections are independent of their subsequent decays, so depend only
on the initial state masses ---
$m(\chargino_1, \neutralino_2)$ for C1N2, and
$m(\neutralino_1)$ for GMSB.

Uncertainties in these cross-sections are assigned from an envelope
(using symmetrized maximum deviations)
over scale variations~\cite{Fuks:2012qx} and parton distribution variations
with the
\textsc{CTEQ6.6}~\cite{Nadolsky:2008cw} and
\textsc{MSTW2008nlo90cl}~\cite{Martin:2009iq} PDF sets.
These uncertainties increase with parent mass from around
$\pm4\%$ at $100\,\eV[G]$ to around
$\pm11\%$ at $800\,\eV[G]$ for both C1N2 and GMSB models~\cite{
atlas_twiki_xsec_c1n2,
atlas_twiki_xsec_hino
}.

\paragraph{Theory variations}
\label{sec:2ljets_theory_variations}
\TODO{put this above alternative samples?}
\TODO{this appears to be signal theory}
% extrapolation
The total theory variation is a combination of seven components, representing
\TODO{what are they?}.
A relative uncertainty due to each, $r$, is evaluated at each benchmark
point $\vec y_i$, and extrapolated elsewhere as an exponentially weighted
combination
\begin{equation}
\label{eqn:2ljets_signal_extrap}
r(\vec x) = \frac{
\sum_i r_i e^{-d(\vec x, \vec y_i)}
}{
\sum_i e^{-d(\vec x, \vec y_i)}
}
\end{equation}
in which $d(\vec x, \vec y)$ is a distance function chosen to encode some
physical understanding into the extrapolation.

Parent mass $m(\chargino_1, \neutralino_2) = m_a$
and mass splitting $m_a - m_b$ are the most
characteristic features in C1N2.
From some experimentation, I found that a Manhattan distance in those features
extrapolated nicely.
Since the distance is dimensionful, it needs to be scaled to a reference value,
which I chose as $100\,\eV[G]$; therefore
\begin{equation}
d_\mathrm{C1N2}(\vec x, \vec y) =
\frac{1}{100\,\eV[G]}
\big(
|m_a^{\vec x} - m_b^{\vec y}|
+
|(m_a^{\vec x} - m_b^{\vec x}) - (m_a^{\vec y} - m_b^{\vec y})|
\big)
.
\end{equation}
To put the Higgs branching fraction $B$ on a similar scale to the parent mass
for GMSB models, its distance measure has a separate scale factor;
\begin{equation}
d_\mathrm{GMSB}(\vec x, \vec y) =
\frac{1}{100\,\eV[G]}
|m_a^{\vec x} - m_b^{\vec y}|
+
5
|B^{\vec x} - B^{\vec y}|
.
\end{equation}
Some example extrapolations according to these rules are displayed in
Figure~\ref{fig:2ljets_signal_sys_extrap}.
Variations tend to be small, around $\approx 20\%$, where and only where
yields are large enough for good statistical precision.

@@

>>>>

\subsubsection{Cross-section variations}
\label{sec:2ljets_xsec_variations}
We scale all signal models to cross-sections calculated with the
\textsc{Resummino} package at NLO accuracy in QCD with NLL
(next-to-leading logarithmic) accuracy in the gaugino pair production of our
signal initial states~\cite{Fuks:2012qx, Fuks:2013vua}.
These cross-sections are independent of their subsequent decays, so depend only
on the initial state masses ---
$m(\chargino_1, \neutralino_2)$ for C1N2, and
$m(\neutralino_1)$ for GMSB.

Uncertainties in these cross-sections are assigned from an envelope
(using symmetrized maximum deviations)
over scale variations and parton distribution variations~\cite{Fuks:2012qx}
with the
\textsc{CTEQ6.6}~\cite{Nadolsky:2008cw} and
\textsc{MSTW2008nlo90cl}~\cite{Martin:2009iq} PDF sets.
These uncertainties increase with parent mass from around
$\pm4\%$ at $100\,\eV[G]$ to around
$\pm11\%$ at $800\,\eV[G]$ for both C1N2 and GMSB models~\cite{
atlas_twiki_xsec_c1n2,
atlas_twiki_xsec_hino
}.

\subsubsection{Theory variations}
\label{sec:2ljets_theory_variations}
Another signal variation is assigned to cover sources of uncertainty from
scale variations ($\mu_R$, $\mu_F$, and the merging scale
$Q_\textrm{cut}^\textrm{CKKW}$)) and modelling of radiations and parton
showers (`eigentune' parameters
\texttt{Var1},
\texttt{Var2},
\texttt{Var3a},
\texttt{Var3b},
\texttt{Var3c}~\cite{ATL-PHYS-PUB-2014-021}).
Following the \atlas\ SUSY group
prescription~\cite{atlas_twiki_susytheoretical},
we consider only a coherent scaling of the scales $\mu_R$ and $\mu_F$ together,
so this theory variation includes a total of seven components.

Scale variations have appeared here twice:
first for cross-section variations and
second for these theory variations.
They therefore contribute both in each fit model and in the varied
cross-sections for varied exclusion contours.
This is questionable, but we shall see here that, thankfully, this scale
double-counting makes very little difference.

\paragraph{Extrapolation}
To evaluate each variation requires a new simulation with different parameters,
so to reduce computational costs we use truth-level only
(through SimpleAnalysis~\cite{simpleanalysis_cern}) and evaluate the
variations at only a handful of benchmark signal models.
From these benchmark signal models, we evaluate a relative uncertainty due
due to each component, and extrapolate those variations from these benchmark
examples to others in the parameter plane.

Extrapolation is performed as follows.
Each benchmark point $i$ has a location in the parameter plane $\vec y_i$, and
for a given source of uncertainty $j$ its relative variation has been evaluated
as $r_{ij}$.
To extrapolate these variations across each parameter plane to another signal
model at $\vec x$, we use an exponentially weighted combination
\begin{equation}
\label{eqn:2ljets_signal_extrap}
r_j(\vec x) = \frac{
\sum_i r_{ij} e^{-d(\vec x, \vec y_i)}
}{
\sum_i e^{-d(\vec x, \vec y_i)}
}
\end{equation}
in which $d(\vec x, \vec y)$ is a distance function chosen to encode some
physical understanding into the extrapolation,
and the sum is over all benchmark points.

Parent mass $m(\chargino_1, \neutralino_2) = m_a$
and mass splitting $m_a - m_b$ are the most
characteristic features in C1N2.
From some experimentation, I found that a Manhattan distance in those features
extrapolated nicely.
Since the distance is dimensionful, it needs to be scaled to a reference value,
which I chose as $100\,\eV[G]$; therefore
\begin{equation}
d_\mathrm{C1N2}(\vec x, \vec y) =
\frac{1}{100\,\eV[G]}
\big(
|m_a^{\vec x} - m_b^{\vec y}|
+
|(m_a^{\vec x} - m_b^{\vec x}) - (m_a^{\vec y} - m_b^{\vec y})|
\big)
.
\end{equation}
To put the Higgs branching fraction $B$ on a similar scale to the parent mass
for GMSB models, its distance measure has a separate scale factor;
\begin{equation}
d_\mathrm{GMSB}(\vec x, \vec y) =
\frac{1}{100\,\eV[G]}
|m_a^{\vec x} - m_b^{\vec y}|
+
5
|B^{\vec x} - B^{\vec y}|
.
\end{equation}
Some example extrapolations according to these rules are displayed in
Figure~\ref{fig:2ljets_signal_sys_extrap}.
This figure illustrates the common theme that our theory variations are small
if and only if signal yields are large enough for good statical precision.

@@

\paragraph{Evaluation}
Statistical precision has limited the utility of most variations described in
this chapter.
These are no different.
To reduce statistical noise here, I generated truth samples $10$~times as large
as our central, reconstructed samples.
These sample sizes vary roughly with cross-section from $300\,000$ to
$2.8$~million events per sample, and due to the $\sqrt{n}$ scaling of
Monte Carlo errors, meant approximately $\sqrt{10}\approx 3.2$ times reduced
noise.
However, this remains insufficient since each resulting variation is a
sum in quadrature of variations in $7$ pairs of alternative samples
whose combination will inflate the uncertainty by roughly a factor of
$\sqrt{7}$.
Where signal models' yields has significant in a signal region,
each variation tends to be around $\approx 20\%$ relative error,
and are compatible with the central samples within statistical uncertainties.

Yet another unique sample-comparison prescription now assigns variations as
half of the absolute difference between up and down yields, irrespective of the
central yield.
This differs from the methods previously described in
Sections~\ref{sec:2ljets_zjets_alt} and~\ref{sec:2ljets_diboson_alt},
and is used for consistency with other $\twoljets$ work.

We evaluate and extrapolate each source of uncertainty individually,
and combine all sources by summation in quadrature in each region to form a
single total theory variation~\cite{atlas_twiki_susytheoretical}.
This sum loses all sense of direction, so we choose to give up and down
variations the same $+$ and $-$ signs in all bins.
Since the theory variation parameter is coupled across all bins, it therefore
implements a form of non-uniform normalization factor.

Why the double counting of scale variations does not matter is now apparent:
this variation is dominated by statistical noise, and the inclusion of one
extra source for the scale variations simply scales that noise up by
a factor of $\approx \sqrt{7/6}$.

\emph{%
Special thanks to Sara~Alderweireldt for sharing \textsc{SignalSystTools},
whose excellent scripts were adapted to simulate these signal
samples, and thanks to Jon~Long for contributing to this study.
}
